---
title: "Getting started course tools"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Rstudio

Rstudio will be the primary data analysis and visualization software that you will use for the course
To install follow the instructions:

<https://www.rstudio.com/products/rstudio/download/>.

# Github and Rstudio

To link Rstudio and our version control software (Git), you need to first create a GitHub account, and then configure your computer to understand/interact with git
and then finally link it with R studio

To do this: follow these instructions - get as far as creating a new project - for now leave the project empty; but make sure you create with a git repository

<https://www.r-bloggers.com/rstudio-and-github/>

If you are using a newer version of MacOS, you may also need to install X-quartz so that R can open graphics windows on the fly

<https://www.xquartz.org/>

You will need to re-start R 

# Install Packages
load some libraries that are likely to be useful

```{r install-packages, eval=F}
# Run this chunk only once in your Console
# Do not evaluate when knitting Rmarkdown

# list of packages
pkgs = c(
  'readr',        # read csv
  'readxl',       # read xls
  'dplyr',        # data frame manipulation
  'tidyr',        # data tidying
  'ggplot2')      # graphics package
    

# install packages if not found
for (p in pkgs){
  if (!require(p, character.only=T)){
    install.packages(p)
  
  }
}
```


# Simple example
You will first need to download clim.txt from ; and move it to your working directory - the directory you created when you made your first project in Rstudio. Now use the following R code to read that file in to R and do some basic analysis
You man need to set your working directory to the project directory

```{r readingexample, include=TRUE}
library(dplyr)
clim = read.table("clim.txt", sep=",", header=T)
# generate R understandable dates
clim$date = as.Date(paste(clim$day, clim$month, clim$year, sep="/"),"%d/%m/%Y")

# look at the file
View(clim)
# see some summary statistics of the columns
summary(clim)
# create a new table that summarizes by wateryear, use mean for states, and sum for fluxes
a = clim %>% select(wy,month,tmax, tmin) %>% group_by(wy, month) %>% 
    summarize_each(funs(mean))
   
b = clim %>% select(wy,month,rain) %>% group_by(wy, month) %>% 
    summarize_each(funs(sum))
   
clim.mwy = inner_join(a,b)
View(clim.mwy)
  
tmp = clim

View(clim.mwy)

#add a new variable
clim$tavg = (clim$tmin+clim$tmax)/2
range(clim$tavg)
range(clim$tmin)
range(clim$tmax)


```

You have 3 options for writing R code

*  directly type into the console (temporary)
*  open an R script file (can be saved to git)
*  **add as a R module in an R Markdown file** (can be saved to git)

## Including Plots



```{r plots, echo=TRUE}
# plots
#simple time series
library(ggplot2)
ggplot(clim, aes(x=date,y=tmax))+geom_line()

#simple boxplot
ggplot(clim, aes(x=as.factor(month), y=tmax))+geom_boxplot()+labs(x="Month",y="Max T (C)")

# use ggplot to aggregate for you
g=ggplot(clim, aes(x=as.numeric(wy), y=tmax))
g= g+stat_summary(fun.y=mean, geom="line", col="red")
g 

# add some prettier axis
g= g+scale_x_continuous(breaks = seq(from=1940,to=2020,by=10), limits=c(1940,2020))

# add another line
g= g+ stat_summary(aes(x=as.numeric(wy), y=tmin), geom="line", col="blue", fun.y=mean)

# add some labels
g = g+labs(x="Water Year", y="Temperature (C)")
g

# subsetting 'on the fly'
ggplot(subset(clim, clim$wy < 1950), aes(x=as.factor(wy), y=rain))+stat_summary(fun.y=sum, geom="bar", col="green", fill="blue")+labs(x="Water Year",y="Rain (mm)")


```

Note that in markdown the `echo = FALSE` parameter can be added to the code chunk to prevent printing of the R code that generated the plot.

```{r plotsnoecho, echo=FALSE}

ggplot(subset(clim, clim$wy < 1950), aes(x=as.factor(wy), y=rain))+stat_summary(fun.y=sum, geom="bar", col="green", fill="blue")+labs(x="Water Year",y="Rain (mm)")
```

# Git steps

Look at Git - 
## Tracking things locally with git

* Check on the file you want to stage
* Commit
     * This will create your master branch
* Make a change to a file
* Use diff to see what changed 
     * you must save the file first
* Commit changes
* Use history to see how things have changed 
* Repeat as you change things

## Saving to your global github (for sharing)

* Make sure you have committed all recent changes
* get the http address from your github repo
* Open a shell in Git tab in Rstudio

type

git remote add origin **address**

git push -u origin master

* Notice how now under the Git tab you have colored arrows, you can now use these to push and pull from github to your local repo
