---
title: "writing.submodels"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Running functions

You can both create your own functions and utilize functions that other people have written - re-use!
User contributed software is available in packages - three useful packages for ecohydrology are
* **Evapotranspiration**
* **EcoHydRology**


the following website provides some additional tools:
[link] (http://abouthydrology.blogspot.co.il/2012/08/r-resources-for-hydrologists.html)
There are many others and new packages are being developed all the time but these are a useful place to start

```{r install-packages, eval=T, echo=F}
# Run this chunk only once in your Console
# Do not evaluate when knitting Rmarkdown

# list of packages
pkgs = c(
  'dplyr',        # data frame manipulation
  'tidyr',        # data tidying
  'Evapotranspiration',   # many evapotranspiration submodels
  'EcoHydRology',  # energy budget and ET
  'ggplot2')      # graphics package
    

# install packages if not found
for (p in pkgs){
  if (!require(p, character.only=T)){
    install.packages(p)
  
  }
}
library(ggplot2)
library(Evapotranspiration)
library(EcoHydRology)

```

## Examples

You will need to read the help material on these packages to both the functions that are available and the inputs that they require
let look at  few examples

```{r userexamples, echo=TRUE}

#start with a simple function - Longwave Radiation a body with a given temperature
# look at function
Longwave

# estimate longwave from atmosphere for a sequence of air temperature
# use our climate data
clim = read.table("data/clim.txt", header=T, sep=",")

# lets start with an atmospheric emissivity of around 0.7
# reasonable value for clear sky and look at nightime longwave
result = Longwave(0.7, clim$tmin)
plot(result)
# add to our clim data frame to make more interesting reulst
clim$nightclearLW = result
ggplot(clim, aes(y=nightclearLW, x=as.factor(month))) +geom_boxplot() 
# but we know that atmospheric emissivity is much higher when it is cloudy - so lets compute an emmissivity
clim$emissivity = ifelse(clim$rain > 0, 0.9, 0.7)

clim$nightLW = Longwave(clim$emissivity, clim$tmin)
p=ggplot(clim, aes(y=nightLW, x=as.factor(month), group=1)) +stat_summary(fun.y=mean, col="blue", geom="line")
p
p+stat_summary(aes(y=nightclearLW, x=as.factor(month), group=1),
               fun.y=mean, col="red", geom="line")
ggplot(clim, aes(y=nightLW-nightclearLW, x=as.factor(month), group=1)) +stat_summary(fun.y=mean, geom="line", col="red")
#
#from the EcoHydRology package - estimation of ET from Priestley-Taylor (PET from temperature)
#lets start by looking at the functio

PET_fromTemp
PTpet

#now lets try with some examples

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
