---
title: "Functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Penman function
```{r functions, eval=FALSE, echo=FALSE}
source("penman_montieth.R")
penman_montieth

#get documentation to work
# include devtools and roxygen2 packages
library("devtools")
library("roxygen2")

# this will create a package called #ecohydrotools and initialize it with R code #for penman
package.skeleton(list="penman_montieth", name ="ecohydrotools")

# go into ecohydrotools directory
setwd("ecohydrotools")
document()

#this should add files to the man subdirectory
help(penman_montieth)
```

## Adding Data

you can also store files in R format in
the data directory - 
**.RData** and **.rda** are standard extensions

and **data** is the standard directory used



```{r data, echo=TRUE}

setwd("../")
getwd()
metdata = read.table("data/metdata.txt", header=T)
metdata$date = as.Date(paste(metdata$year, metdata$month, metdata$day, sep="/"))

# one can save metdata in R format to the package
dir.create("ecohydrotools/data")
save(metdata, file="ecohydrotools/data/metdata.RData")
```
## Applying
Now make use of your code and data
If re-open Rstudio, you won't have these files in your workspace
The easiest way to get them in is to use **source** and then repeat documentation

Best though trickier is to actually implement your package

```{r apply, echo=TRUE}
# use data in functions in ecohydrotools package

setwd("./ecohydrotools")
source("R/penman_montieth.R")
document()
# getting data
data("metdata")
etres=with(metdata,penman_montieth(Tair=tavg, vpd=vpd, Rnet=rnet, gs=gs, ga=ga, dayl=24))

install.packages("ecohydrotools")
library("ecohydrotools")



etres=with(metdata,penman_montieth(Tair=tavg, vpd=vpd, Rnet=rnet, gs=gs, ga=ga, dayl=24))

#add to data set - so that we can plot variables
metdata$et = etres

plot(metdata$date, metdata$et, type="l", col="blue", 
     ylab="ET in mm/day", xlab="Date")

# imagine gs is from a larger tree double the LAI

etres=with(metdata,penman_montieth(Tair=tavg, vpd=vpd, Rnet=rnet, gs=gs*4, ga=ga, dayl=24))
lines(metdata$date, etres, col="green")

# monthly values
with(metdata, boxplot(et~month, ylab="ET (mm/day)", xlab="Month"))

```
