---
title: "Functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Documentation

We can automatically create documentation with the Penman function as an example

```{r functions, eval=FALSE, echo=TRUE}
source("penman_montieth.R")
penman_montieth

#get documentation to work
# include devtools and roxygen2 packages
library("devtools")
library("roxygen2")

# this will create a package called #ecohydrotools and initialize it with R code for penman_montieth
package.skeleton(list="penman_montieth", name ="ecohydrotools")

# go into ecohydrotools directory
setwd("ecohydrotools")
document()

#this should add files to the man subdirectory
help(penman_montieth)
```

## Adding Data

you can also store files in R format in
the data directory - 
**.RData** and **.rda** are standard extensions

and **data** is the standard directory used



```{r data, echo=TRUE, eval=FALSE}

setwd("/Users/naomi/UCSB.GoogleDrive/courses/Israel/Recohydro/")
metdata = read.table("data/metdata.txt", header=T)
metdata$date = as.Date(paste(metdata$year, metdata$month, metdata$day, sep="/"))

# one can save metdata in R format to the package
dir.create("ecohydrotools/data")
save(metdata, file="ecohydrotools/data/metdata.RData")
```
## Applying
Now make use of your code and data

If re-open Rstudio, you won't have these files in your workspace

The easiest way to get them in is to use **source** and then repeat documentation

Best though trickier is to actually implement your package

```{r apply, echo=TRUE}
# use data in functions in ecohydrotools package

setwd("/Users/naomi/UCSB.GoogleDrive/courses/Israel/Recohydro/ecohydrotools")
source("R/penman_montieth.R")
library(roxygen2)
library(devtools)
document()
# getting data
data("metdata")
etres=with(metdata,penman_montieth(Tair=tavg, vpd=vpd, Rnet=rnet, gs=gs, ga=ga, dayl=24))




#add to data set - so that we can plot variables
metdata$et = etres

plot(metdata$date, metdata$et, type="l", col="blue", 
     ylab="ET in mm/day", xlab="Date")

# imagine gs is from a larger tree double the LAI

etres=with(metdata,penman_montieth(Tair=tavg, vpd=vpd, Rnet=rnet, gs=gs*4, ga=ga, dayl=24))
lines(metdata$date, etres, col="green")

# monthly values
with(metdata, boxplot(et~month, ylab="ET (mm/day)", xlab="Month"))

```

## Making a package 

Alternatively, we can make a package that you re-load (this is called building and reloading)

Go to Tools, Project Options, Build Tools and write in in the directory of the package you created
(check dev.tools and roxygen)

Then in Build (on Rstudio options)
Load all
and when you make changes
Build and Reload

When you start Rstudio with your project; after having closed your session
go to Build; load all
This will add everything

```{r loadpackage, echo=TRUE}

data(metdata)
etres=with(metdata,penman_montieth(Tair=tavg, vpd=vpd, Rnet=rnet, gs=gs, ga=ga, dayl=24))
metdata$et = etres

plot(metdata$date, metdata$et, type="l", col="blue", 
     ylab="ET in mm/day", xlab="Date")

```

# adding more to your package

Create your .R file with proper documentation format and then add it to the ecohydrotools/R directory

Lets try with compute_yield.R, if its in that directory 
document will automoatically find it
```{r adaptpackage, echo=TRUE}

setwd("/Users/naomi/UCSB.GoogleDrive/courses/Israel/Recohydro/ecohydrotools")
document()
help(compute_yield)

compute_yield(4, 200, 0.3, 0.4, 5)
# to see code
compute_yield

```

Remember to add and commit these changes to you local git and when you are ready push to github